---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load, cache=TRUE, echo=TRUE}
wdir <- getwd()
if (wdir != "E:/Luciano/R_WorkingDir/Reproducible Research/RepData_PeerAssessment1"){
    setwd("E:/Luciano/R_WorkingDir/Reproducible Research/RepData_PeerAssessment1")
}
rm(list = ls())
listfiles <- dir()
if ("activity.csv" %in% listfiles){
    actdf <- read.csv("activity.csv")
    newdf <- na.omit(actdf)
}else{
    if ("activity.zip" %in% listfiles){
        unzip(zipfile = "activity.zip")
        actdf <- read.csv("activity.csv") 
        newdf <- na.omit(actdf)
    }else{
        print("No files found!!!")
    }
}
```


## What is mean total number of steps taken per day?
- Update total steps per day.
- Show histogram with totals steps per day.
- Calculate mean and median of steps per day
```{r hist, echo=TRUE}
instpackages <- installed.packages()
if (("ggplot2" %in% instpackages)){
    library(ggplot2)
}else{
    install.packages("ggplot2")
    library(ggplot2)
}
totaldays <- tapply(newdf$steps, newdf$date, FUN = sum)
hist(totaldays, main = "Histogram of the total number of steps", 
     xlab = "Total number of steps taken each day", 
     col = "grey", border = "green")
meansteps <- mean(totaldays, na.rm = TRUE)
mediansteps <- median(totaldays, na.rm = TRUE)
abline(v = meansteps, col = "royalblue", lwd = 2)
abline(v = mediansteps, col = "red", lwd = 2)
legend("topright", c("- Median", "- Mean"), fill = c("royalblue", "red"))
box()
```

- **Mean: ** `r meansteps`
- **Median: ** `r mediansteps`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
